What We Eval in the Shadows

### 2021-04-05 paper/main.tex
Removed the legacy text

\paragraph{Base}
The spread of forms in Base are similar to CRAN with a larger share of \c{V}, as
it is present in \baseMinimizedpropsitesa\% of the sites and \c{X} is a distant
second with \baseMinimizedpropsitesb\%.

\paragraph{Kaggle}
Kaggle is dominated by \c{V}, with \kaggleMinimizedpropsitesa\% of the
\kaggleNbCallSites sites.

@@

\section{LEGACY TEXT}

We looked at the top ten expressions passed to core and package \eval
calls. The most frequent ten expressions to eval calls from core R
contribute to 85\% of all \eval calls.

\begin{table}[!h] \centering
\begin{tabular}{@{}l|rr@{}} \hline
Expression & Eval Call &  \% \\\hline
\c{c("auto", "shell", "radix")} & 1,987,105 & 29\%\\
\c{c("auto", "shell", "quick", "radix")} & 1,593,169  & 23\%\\
\c{\{info <- loadingNamespaceInfo(...} & 1,008,632 &       14\%\\
\c{c("onLoad", "attach", "detach", "onUnload")}   & 470,566 &      6.9\%\\
\c{c("append", "prepend", "replace")} &              261,587&       3.9\% \\
\c{c("left", "right", "centre", "none")} & 162,086     & 2.4\%\\
\c{c("no", "ifany", "always")}   &                71,580 &       1.1\%\\
\c{c("pearson", "kendall", "spearman")}  & 72,962 &      1.1\%\\
\c{NULL}& 75,330  &      1.1\% \\
\c{Symbol}&                 66,279&       1\%\\\hline
\end{tabular}
\caption{Top ten eval calls in Core}\label{B}
\end{table}

The expression \c{\{info <- loadingNamespaceInfo(...} is added by core R to a
package directory during installation. To load the package, this code is
executed. It creates a namespace for the package, injects the package bindings,
and attaches the namespace to the program search path. The \c{NULL} comes from a
call to \c{substitute(subset)} in \c{stats::model.frame.default} function which
has a default value of \c{subset} as \c{NULL}. The \c{Symbol} arises from a call
to \c{as.name} in \c{base::str} function that returns a symbol that is looked up
by evaluating it in a specific environment. The remaining cases arise from calls
to \c{match.arg} which is used to look up the default choices for a variable and
match against the choice passed by the caller.

The most frequent ten expressions to eval calls from CRAN packages
contribute to 77.1\% of all \eval calls.

\begin{table}[!h]  \centering
\begin{tabular}{@{}l@{~}|@{~}r@{~}r@{}} \hline
Expression & Eval Call &  \% \\\hline
\c{Environment} &                                  989302   & 61\%\\
\c{column[rows] <<- what} &                        55677    & 3.5\%\\
\c{function(value) freduce(value, `function.list`)} & 37251& 2.3\%\\
\c{NULL} &                         32005    & 2\%\\
\c{List} &                         22293    & 1.4\%\\
\c{c("default", "default2012", "default2011" ...}& 20610    & 1.3\%\\
\c{force(..1)}            &                        20461    & 1.3\%\\
\c{alist(`spec`)}       &                         18532    & 1.2\%\\
\c{inner}               &                          18530    & 1.2\%\\
\c{String Vector}      &                           17487     & 1.1\%\\
\end{tabular}\caption{Top ten eval calls in CRAN} \label{C}
\end{table}

The expression \c{Environment} occurs because of the four callsites explained
above, \c{ggplot2::ggproto}, \c{R6::generator\_funs}, \c{future::backtrace} and
\c{RModel::str.RMmodel}. The next expression, \c{column[rows] <<- what}, is used
inside the \c{plyr::rbind.fill} function to merge data frames by assigning
concatenated vectors to rows. The \c{<<-} operator is interesting in that it
skips the current scope and assigns in a parent scope in which the variable is
already present. In our corpus, all these \eval calls contribute to a single
side-effect. The expression \c{function(value) freduce(value, `function.list`)}
arises from the \c{magrittr::\%>\%} function which is a pipe operator that pipes
the output of previous command to the next one. The expression is evaluated in a
custom environment to create a function binding for evaluating the components of
the pipe.\c{String Vector} and \c{List} also arise from the same function when a
string or a list is piped using the \c{\%>\%} function into the next expression.
The \c{NULL} arises from \c{R6::generator\_funs} function when the \eval is
passed a \c{NULL} argument by the \c{DataMask\_generator} package. The
\c{c("default", "default2012", "default2011" ...} pattern arises from
\c{copula::polyG} where it reflectively access the default expression for its
formal parameter and evaluates it. The \c{force(..1)} and \c{alist(`spec`)}
patterns occur in \c{glue::glue\_data} function which concatenates and
interpolates strings. The two patterns occur because the function captures
unevaluated unnamed arguments and maps the evaluation of \c{force(..1)} on them.
The \c{force} function forces promises and returns the result of evaluation. The
\c{inner} pattern arises from \c{glue::identity\_transformer} which enables the
creation of custom transformation functions for affecting the interpolation and
concatenation of input by the \c{glue} package.

\section{Why do we Eval}

\subsection{Discussion}

Yeah, why?

\subsection{Can we do without?}

We look at how \emph{consistent} the \c{expr} argument of \eval can be, \ie how
many different types of the resolved \c{expr} there are per call sites. Most of
the call sites, \ie \PercentMonomorphic, are \emph{consistent}, and this is
similar to javascript. However, a few ones are highly \emph{polymorphic} (10
different types). They are the pipe operators \c{\%>\%}, \c{\%<>\%} and
\c{\%\$\%} in package \emph{magrittr}. It is effectively used to compose
functions on their first argument, which can be of any type.

@@

>>>>

\paragraph{Base}
\todo{I would remove this - we conclude Base in 4}
The spread of forms in Base are similar to CRAN with a larger share of \c{V}, as
it is present in \baseMinimizedpropsitesa\% of the sites and \c{X} is a distant
second with \baseMinimizedpropsitesb\%.

\paragraph{Kaggle}
\todo{I would remove this - we conclude Kaggle in 4}
Kaggle is dominated by \c{V}, with \kaggleMinimizedpropsitesa\% of the
\kaggleNbCallSites sites.

\todo{The following should be either removed or positioned properly}
We look at how \emph{consistent} the \c{expr} argument of \eval can be, \ie how
many different types of the resolved \c{expr} there are per call sites. Most of
the call sites, \ie \PercentMonomorphic, are \emph{consistent}, and this is
similar to javascript. However, a few ones are highly \emph{polymorphic} (10
different types). They are the pipe operators \c{\%>\%}, \c{\%<>\%} and
\c{\%\$\%} in package \emph{magrittr}. It is effectively used to compose
functions on their first argument, which can be of any type.

@@

\mypara{Unnecessary use of \eval}

@@

\todo{Aviral's story when he contacted a developer about his use of eval}
